---test %>%
 group_by(Type) %>%
  select(Type)
title: "PA Hackathon - Home Prices"
author: "Christopher Chen"
date: '2022-05-13'
output: html_document
---


```{r }
library(readr)
library(caret)
library(dplyr)
library(corrplot)
library(RColorBrewer)
library(tidyr)
library(MASS)
library(tree)
library(randomForest)
```


```{r }
train <- read_csv("C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/beginner_train - Original.csv")
test <- read_csv("C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/beginner_test - Original.csv")
```


```{r } 
#Outliers Handling 
train_new <- train %>%
  	mutate(
  	  Total_spaces = ifelse(Total_spaces<0, abs(Total_spaces), Total_spaces),
  	  Garage_spaces = ifelse(Garage_spaces<0, abs(Garage_spaces), Garage_spaces),
  	  Tax_assessed_value = ifelse(Listed_Price==0,0,Tax_assessed_value),
  	  Annual_tax_amount = ifelse(Listed_Price==0,0,Annual_tax_amount),
  	  Bedrooms_N = as.numeric(Bedrooms),
  	  Bathrooms_N = as.numeric(Bathrooms),
  	  Bedrooms_N1 = ifelse(is.na(Bedrooms_N),as.numeric(Bathrooms),Bedrooms_N),
  	  Bathrooms = ifelse(Bathrooms_N == 0, NA, Bathrooms_N),
  	  Bedrooms = ifelse(Bedrooms_N1 == 0, NA, Bedrooms_N1),
  	  Total_spaces = ifelse(Total_spaces == 0, NA, Total_spaces),
  	  Garage_spaces = ifelse(Garage_spaces == 0, NA, Garage_spaces),
  	  Total_interior_livable_area = ifelse(Total_interior_livable_area>50000,NA,Total_interior_livable_area),
  	  Lot = ifelse(Lot>100000000 | Lot <= 1,NA,Lot),
  	  Last_Sold_Price = ifelse(Last_Sold_Price<1000,0,Last_Sold_Price),
  	  Listed_On = as.Date(Listed_On, format = "%m/%d/%Y"),
  	  Last_Sold_On = ifelse(is.na(Last_Sold_On), NA, as.Date(Last_Sold_On, format = "%m/%d/%Y")),
  	  Listed_On = ifelse(Listed_On < Last_Sold_On & !is.na(Last_Sold_On),Last_Sold_On,Listed_On),
    	Type = ifelse(Type %in% c("SingleFamily","Condo","Townhouse", "MultiFamily","MobileManufactured","VacantLand"), Type, "Unknown"),
  	  Listed_Price = ifelse(Last_Sold_Price>Listed_Price, Last_Sold_Price, Listed_Price),
  	  Sold_Price = ifelse(Id==19827, Listed_Price,Sold_Price),
  	  Listed_Price = ifelse(((Sold_Price-Listed_Price)/Listed_Price) >= 1, Sold_Price, Listed_Price),
  	  Year_built = ifelse(Id==3180, NA, Year_built),
  	  Last_Sold_Price = ifelse(is.na(Last_Sold_Price),Tax_assessed_value,Last_Sold_Price))
  


```



```{r }
#Auto fill for missing values
train_new1 <- train_new %>%
  filter(!(Year_built < 1000 | Year_built > 2200)) %>%
  arrange(Sold_Price, Lot) %>%
  fill(Lot, .direction = 'down') %>%
  arrange(Sold_Price, Full_bathrooms) %>%
  fill(Full_bathrooms, .direction = 'down') %>%
  arrange(Sold_Price, Bathrooms) %>%
  fill(Bathrooms, .direction = 'down') %>%
  arrange(Sold_Price, Bedrooms) %>%
  fill(Bedrooms, .direction = 'down') %>%
  arrange(Sold_Price, Total_interior_livable_area) %>%
  fill(Total_interior_livable_area, .direction = 'down') %>%
  arrange(Sold_Price, Total_spaces) %>%
  fill(Total_spaces, .direction = 'down') %>%
  arrange(Sold_Price, Garage_spaces) %>%
  fill(Garage_spaces, .direction = 'down') %>%
  arrange(Sold_Price, Elementary_School_Score) %>%
  fill(Elementary_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, Elementary_School_Distance ) %>%
  fill(Elementary_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, Middle_School_Score) %>%
  fill(Middle_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, Middle_School_Distance) %>%
  fill(Middle_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, High_School_Score) %>%
  fill(High_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, High_School_Distance) %>%
  fill(High_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, Listed_Price) %>%
  fill(Listed_Price, .direction = 'down') %>%
  arrange(Sold_Price, Last_Sold_Price) %>%
  fill(Last_Sold_Price, .direction = 'down') %>%
  arrange(Sold_Price, Tax_assessed_value) %>%
  fill(Tax_assessed_value, .direction = 'down') %>%
  arrange(Sold_Price, Annual_tax_amount) %>%
  fill(Annual_tax_amount, .direction = 'down') %>%
  filter(!(is.na(Garage_spaces)))
  

# create normalization function
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
# create normalization function
factorize <- function(x) {
  return (as.factor(x))
}
# normalize the DataSet data
#DataSet_n <- as.data.frame(lapply(train_new1[,1:20], factorize))
DataSet_n <- train_new1 %>%
            mutate(Time_Diff = as.Date("5/15/2022", format = "%m/%d/%Y") - Listed_On,
  	               Inflation = (Sold_Price - Listed_Price) / as.numeric(Time_Diff))


```


```{r }
#test transformed
test_new <- test %>%
            mutate(Sold_Price = Listed_Price,
                    Bedrooms_N = as.numeric(Bedrooms),
          	        Bathrooms_N = as.numeric(Bathrooms),
          	        Bedrooms_N1 = ifelse(is.na(Bedrooms_N),as.numeric(Bathrooms),Bedrooms_N),
          	        Bathrooms = ifelse(Bathrooms_N == 0, NA, Bathrooms_N),
          	        Bedrooms = ifelse(Bedrooms_N1 == 0, NA, Bedrooms_N1),
          	    	  Listed_On = as.numeric(as.Date(Listed_On, format = "%m/%d/%Y")),
          	        Last_Sold_On = ifelse(is.na(Last_Sold_On), NA, as.Date(Last_Sold_On, format = "%m/%d/%Y")))

#train_new1 = train_new1[,!(names(train_new1) %in% c("Bedrooms_N","Bathrooms_N","Bedrooms_N1"))]
#test_new = test_new[,!(names(test_new) %in% c("Bedrooms_N","Bathrooms_N","Bedrooms_N1"))]
#test_new3 <- union(train_new1,test_new)

test_new <- test_new %>%
mutate(
  	  Total_spaces = ifelse(Total_spaces<0, abs(Total_spaces), Total_spaces),
  	  Garage_spaces = ifelse(Garage_spaces<0, abs(Garage_spaces), Garage_spaces),
  	  Tax_assessed_value = ifelse(Listed_Price==0,0,Tax_assessed_value),
  	  Annual_tax_amount = ifelse(Listed_Price==0,0,Annual_tax_amount),
  	  Bedrooms_N = as.numeric(Bedrooms),
  	  Bathrooms_N = as.numeric(Bathrooms),
  	  Bedrooms_N1 = ifelse(is.na(Bedrooms_N),as.numeric(Bathrooms),Bedrooms_N),
  	  Bathrooms = ifelse(Bathrooms_N == 0, NA, Bathrooms_N),
  	  Bedrooms = ifelse(Bedrooms_N1 == 0, NA, Bedrooms_N1),
  	  Total_spaces = ifelse(Total_spaces == 0, NA, Total_spaces),
  	  Garage_spaces = ifelse(Garage_spaces == 0, NA, Garage_spaces),
  	  Total_interior_livable_area = ifelse(Total_interior_livable_area>50000 | Total_interior_livable_area <10,NA,Total_interior_livable_area),
  	  Lot = ifelse(Lot>100000000 | Lot <= 1,NA,Lot),
  	  Last_Sold_Price = ifelse(Last_Sold_Price<1000,0,Last_Sold_Price),
    	Type = ifelse(Type %in% c("SingleFamily","Condo","Townhouse", "MultiFamily","MobileManufactured","VacantLand"), Type, "Unknown"),
  	  Year_built = ifelse(Year_built==0|Year_built>2200,NA,Year_built))
  

```

```{r }
test_new1 <- test_new %>%
  arrange(Sold_Price, Year_built) %>%
  fill(Year_built, .direction = 'down') %>%
  arrange(Sold_Price, Lot) %>%
  fill(Lot, .direction = 'down') %>%
  arrange(Sold_Price, Full_bathrooms) %>%
  fill(Full_bathrooms, .direction = 'down') %>%
  arrange(Sold_Price, Bathrooms) %>%
  fill(Bathrooms, .direction = 'down') %>%
  arrange(Sold_Price, Bedrooms) %>%
  fill(Bedrooms, .direction = 'down') %>%
  arrange(Sold_Price, Total_interior_livable_area) %>%
  fill(Total_interior_livable_area, .direction = 'down') %>%
  arrange(Sold_Price, Total_spaces) %>%
  fill(Total_spaces, .direction = 'down') %>%
  arrange(Sold_Price, Garage_spaces) %>%
  fill(Garage_spaces, .direction = 'down') %>%
  arrange(Sold_Price, Elementary_School_Score) %>%
  fill(Elementary_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, Elementary_School_Distance ) %>%
  fill(Elementary_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, Middle_School_Score) %>%
  fill(Middle_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, Middle_School_Distance) %>%
  fill(Middle_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, High_School_Score) %>%
  fill(High_School_Score, .direction = 'down') %>%
  arrange(Sold_Price, High_School_Distance) %>%
  fill(High_School_Distance, .direction = 'down') %>%
  arrange(Sold_Price, Listed_Price) %>%
  fill(Listed_Price, .direction = 'down') %>%
  arrange(Sold_Price, Last_Sold_Price) %>%
  fill(Last_Sold_Price, .direction = 'down') %>%
  arrange(Sold_Price, Tax_assessed_value) %>%
  fill(Tax_assessed_value, .direction = 'down') %>%
  arrange(Sold_Price, Annual_tax_amount) %>%
  fill(Annual_tax_amount, .direction = 'down')

#write.csv(test_new1,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/Test_new1.csv")

test_new1 <- test_new1 %>%
    mutate(Time_Diff = as.Date("5/15/2022", format = "%m/%d/%Y") - Listed_On,
  	     Inflation = (Sold_Price - Listed_Price) / as.numeric(Time_Diff)) %>%
    filter(Id >= 37439)

test_new1 = test_new1[,!(names(test_new1) %in% c("Listed_On","Last_Sold_On","Bedrooms_N","Bathrooms_N","Bedrooms_N1"))]

#Transformed Data
dmy <- dummyVars(" ~ .", data = test_new1)
transformed_data_test <- data.frame(predict(dmy, newdata = test_new1))
  
```

Approach 1 - train with Regression Tree (low precision)
```{r }
#Fitting with Regression Tree model 
set.seed(1)

train = sample(1:nrow(transformed_data), nrow(transformed_data)/2)
tree.boston=tree(Inflation~.,transformed_data,subset=train)
summary(tree.boston)
plot(tree.boston)
text(tree.boston,pretty=0)

transformed_data_test$yhat=predict(tree.boston,newdata=transformed_data_test)
write.csv(transformed_data_test,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/Test_new_tree1.csv")
```

Approach 1 - train with Random Forest (Best accuracy model)
```{r }
#Fitting with Random Forest (Best accuracy)
set.seed(1)
dmy <- dummyVars(" ~ .", data = DataSet_n)
transformed_data <- data.frame(predict(dmy, newdata = DataSet_n))

transformed_data = transformed_data[,(names(transformed_data) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","TypeSingleFamily","Inflation","Time_Diff"))]

transformed_data_test_train = transformed_data_test[,(names(transformed_data_test) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","TypeSingleFamily","Inflation","Time_Diff"))]

transformed_data_train <- union(transformed_data,transformed_data_test_train)


bag.boston=randomForest(Sold_Price~.,data=transformed_data_train,mtry=14,importance=TRUE)
bag.boston

transformed_data_test_2 = transformed_data_test[,(names(transformed_data_test) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Id","Elementary_School_Score","Middle_School_Score","High_School_Score","TypeSingleFamily","Inflation","Time_Diff"))]



transformed_data_test_2$yhat=predict(bag.boston,newdata=transformed_data_test_2)

write.csv(transformed_data_test_2,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/Test_new_bag6.csv")


```

Approach 1 - train with GBM (Medium accuracy model with negative prediction value)
```{r }
#Fitting with GBM
library(gbm)
set.seed(1)
dmy <- dummyVars(" ~ .", data = DataSet_n)
transformed_data <- data.frame(predict(dmy, newdata = DataSet_n))

transformed_data = transformed_data[,(names(transformed_data) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Inflation","Time_Diff"))]

transformed_data_test_train = transformed_data_test[,(names(transformed_data_test) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Inflation","Time_Diff"))]

transformed_data_train <- union(transformed_data,transformed_data_test_train)

boost.boston=gbm(Sold_Price~.,data=transformed_data_train,distribution="gaussian",n.trees=10000,interaction.depth=4)


transformed_data_test_2 = transformed_data_test[,(names(transformed_data_test) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Inflation","Time_Diff","Id"))]
transformed_data_test_2$yhat.boost=predict(boost.boston,newdata=transformed_data_test_2,n.trees=10000)

write.csv(transformed_data_test_2,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/test_new1_boost7.csv")

plot(transformed_data_test$Listed_Price, transformed_data_test$yhat.boost)
abline(0,1)
```
Correlations Plot 
```{r }
# create normalization function
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# normalize the DataSet data

M<-cor(train_new2)
corrplot(M, method="color")

```




#Approach 2 -
Consolidate Train + Test, missing values are assigned with -1 and kept in set.
```{r  } 
train_new2 <- train_new %>%
  mutate(Bedrooms = ifelse(is.na(Bedrooms),-1,Bedrooms),
         Bathrooms = ifelse(is.na(Bathrooms),-1,Bathrooms),
         Full_bathrooms = ifelse(is.na(Full_bathrooms),-1,Full_bathrooms),
         Total_interior_livable_area = ifelse(is.na( Total_interior_livable_area),-1,Total_interior_livable_area),
         Tax_assessed_value = ifelse(is.na(Tax_assessed_value),-1,Tax_assessed_value),
         Annual_tax_amount = ifelse(is.na(Annual_tax_amount),-1,Annual_tax_amount),
         Listed_Price = ifelse(is.na(Listed_Price),-1,Listed_Price),
         Last_Sold_Price = ifelse(is.na(Last_Sold_Price),-1,Last_Sold_Price),
         Elementary_School_Score = ifelse(is.na(Elementary_School_Score),-1,Elementary_School_Score),
         Middle_School_Score = ifelse(is.na(Middle_School_Score),-1,Middle_School_Score),
         High_School_Score = ifelse(is.na(High_School_Score),-1,High_School_Score),
         Elementary_School_Distance = ifelse(is.na(Elementary_School_Distance),-1,Elementary_School_Distance),
         Middle_School_Distance = ifelse(is.na(Middle_School_Distance),-1,Middle_School_Distance),
         High_School_Distance = ifelse(is.na(High_School_Distance),-1,High_School_Distance),
         Lot = ifelse(is.na(Lot),-1,Lot))


test_new2 <- test_new %>%
  mutate(Bedrooms = ifelse(is.na(Bedrooms),-1,Bedrooms),
         Bathrooms = ifelse(is.na(Bathrooms),-1,Bathrooms),
         Full_bathrooms = ifelse(is.na(Full_bathrooms),-1,Full_bathrooms),
         Total_interior_livable_area = ifelse(is.na( Total_interior_livable_area),-1,Total_interior_livable_area),
         Tax_assessed_value = ifelse(is.na(Tax_assessed_value),-1,Tax_assessed_value),
         Annual_tax_amount = ifelse(is.na(Annual_tax_amount),-1,Annual_tax_amount),
         Listed_Price = ifelse(is.na(Listed_Price),-1,Listed_Price),
         Last_Sold_Price = ifelse(is.na(Last_Sold_Price),-1,Last_Sold_Price),
         Elementary_School_Score = ifelse(is.na(Elementary_School_Score),-1,Elementary_School_Score),
         Middle_School_Score = ifelse(is.na(Middle_School_Score),-1,Middle_School_Score),
         High_School_Score = ifelse(is.na(High_School_Score),-1,High_School_Score),
         Elementary_School_Distance = ifelse(is.na(Elementary_School_Distance),-1,Elementary_School_Distance),
         Middle_School_Distance = ifelse(is.na(Middle_School_Distance),-1,Middle_School_Distance),
         High_School_Distance = ifelse(is.na(High_School_Distance),-1,High_School_Distance),
         Lot = ifelse(is.na(Lot),-1,Lot))
         


```

Approach 2 - Train with (Train + Test) Random Forest 
```{r }
set.seed(1)

train_new2 = train_new2[,(names(train_new2) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Elementary_School_Distance","Middle_School_Distance","High_School_Distance","Lot"))]

test_new2_train = test_new2[,(names(test_new2) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Elementary_School_Distance","Middle_School_Distance","High_School_Distance","Lot"))]

transformed_data_train <- union(test_new2_train,train_new2)

bag.boston=randomForest(Sold_Price~.,data=transformed_data_train,mtry=15,importance=TRUE)
bag.boston

test_new2_test = test_new2[,(names(test_new2) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Id","Elementary_School_Score","Middle_School_Score","High_School_Score","Elementary_School_Distance","Middle_School_Distance","High_School_Distance","Lot"))]

test_new2_test$yhat=predict(bag.boston,newdata=test_new2_test)

test_new2_test <- test_new2_test %>%
    filter(Id >= 37439)

write.csv(test_new2_test,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/Test_new_bag8.csv")


```

Approach 2 - Train with (Train + Test) Random Forest with Test Sold Prices being predicted by the chuck prior - Second Train 
```{r }
test_new2_test <- test_new2_test %>%
  mutate(Sold_Price = yhat)

test_new2_test = test_new2_test[,(names(test_new2_test) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Sold_Price","Elementary_School_Score","Middle_School_Score","High_School_Score","Elementary_School_Distance","Middle_School_Distance","High_School_Distance","Lot"))]

transformed_data_train <- union(test_new2_test,train_new2)

bag.boston=randomForest(Sold_Price~.,data=transformed_data_train,mtry=15,importance=TRUE)

test_new2_test = test_new2[,(names(test_new2) %in% c("Bedrooms","Bathrooms","Full_bathrooms","Total_interior_livable_area","Tax_assessed_value","Annual_tax_amount","Listed_Price","Last_Sold_Price","Id","Elementary_School_Score","Middle_School_Score","High_School_Score","Elementary_School_Distance","Middle_School_Distance","High_School_Distance","Lot"))]

test_new2_test$yhat=predict(bag.boston,newdata=test_new2_test)

test_new2_test <- test_new2_test %>%
    filter(Id >= 37439)

write.csv(test_new2_test,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/Test_new_bag9.csv")

```

Approach 2 - Train with (Train + Test) GBM
```{r }
library(gbm)
set.seed(1)

boost.boston=gbm(Sold_Price~.,data=transformed_data_train,distribution="gaussian",n.trees=10000,interaction.depth=5)

test_new2_test$yhat.boost=predict(boost.boston,newdata=test_new2_test,n.trees=10000)

write.csv(test_new2_test,"C:/Users/chris/Documents/Projects/GitHub R/PA - Hackathon - 20220513/test_new1_boost6.csv")

plot(transformed_data_test$Listed_Price, transformed_data_test$yhat.boost)
abline(0,1)
```


